---
title: "Laboratory 2"
author: "Catherine Vanaise"
date: "1/25/2021"
output: html_document
---

```{r}
library(gsheet)
dataUnido <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1uLaXke-KPN28-ESPPoihk8TiXVWp5xuNGHW7w7yqLCc/edit#gid=416085055")
```

# Piping

```{r}
library(dplyr)
dataUnido %>%
  glimpse()
```

# Loading the dataset into a variable named gdpCountry

```{r}
library(readr)
gdpCountry <- read_csv("./data/gdpCountry.csv")
```

# Loading the dataset into a variable named myData

```{r}
myData <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1uLaXke-KPN28-ESPPoihk8TiXVWp5xuNGHW7w7yqLCc/edit#gid=416085055")
```

# Making a new column

```{r}
dataUnido$newColumn <- 42
```

# Show the columns' name

```{r}
colnames(dataUnido)
```

# Show the column newColumn

```{r}
head(dataUnido[,"newColumn"])
```

# Multiply by 2 and add 5

```{r}
dataUnido$newColumn <- dataUnido$newColumn * 2 + 5
```

# Show the column 'newColumn'

```{r}
head(dataUnido[,"newColumn"])
```

# Rename the newColumn

```{r}
library(dplyr)
dataUnido <- dataUnido %>% 
  dplyr::rename(newColumnRenamed = newColumn)
colnames(dataUnido)
```

# Delete the column named 'newColumn'

```{r}
dataUnido$newColumnRenamed <- NULL
# Show columns' name
colnames(dataUnido)
```

# Erasing non important variables

```{r}
dataUnido$isicCodeCombinaison <- NULL
dataUnido$tableDefinitionCode <- NULL
dataUnido$sourceCode <- NULL
dataUnido$updateYear <- NULL
dataUnido$unit <- NULL
```

# Show columns' name

```{r}
colnames(dataUnido)
```

# Provide the dimension of the dataframe

```{r}
dim(dataUnido)
```

# Subset of dataUnido based on countryCode == Canada

```{r}
library(dplyr)
dataUnidoCanada <- filter(dataUnido, countryCode == 124)
```

# First lines of the dataframe

```{r}
head(dataUnidoCanada)
```

# Number of columns & rows

```{r}
dim(dataUnidoCanada)
```

# Subset of dataUnidoCanada based on two variables (number of employees and establishments)

```{r}
dataUnidoCanadaVariables <- filter(dataUnidoCanada, tableCode == 4 | tableCode == 1)
```

# First lines of the dataframe

```{r}
head(dataUnidoCanadaVariables)
```

# Number of columns & rows

```{r}
dim(dataUnidoCanadaVariables)
```

# Subset of dataUnido based on countryCode == Canada

```{r}
dataUnidoCanadaVariablesAfter2009 <- filter(dataUnidoCanadaVariables, year > 2009)
```

# First lines of the dataframe

```{r}
head(dataUnidoCanadaVariablesAfter2009)
```

# Dimension of the dataframe

```{r}
dim(dataUnidoCanadaVariablesAfter2009)
```

# dataSorted will receive the dataframe dataUnidoCanadaVariablesAfter2009 sorted by the column value

```{r}
dataSorted <- arrange(dataUnidoCanadaVariablesAfter2009, value)
```

# first 6 lines of the dataset

```{r}
head(dataSorted)
```

# dataReverse is the opposite of dataSorted, i.e. the first lines will have the highest values

```{r}
dataReverse <- arrange(dataUnidoCanadaVariablesAfter2009, desc(value))
```

# first 6 lines of the dataset

```{r}
head(dataReverse)
```

# Using pivot_wider() to transform a long dataframe into a wide dataframe

```{r}
library(tidyr)
wideData <- dataSorted %>%
  pivot_wider(names_from = isicCode, values_from = value)
head(wideData)
```

# Using pivot_longer() to transform from wide to long data

```{r}
longData <- wideData %>% 
  pivot_longer(!c(year, tableCode, countryCode), names_to = "isicCode", values_to = "value")
```

# Dimension of the dataframe

```{r}
dim(longData)
```

# First 6 lines

```{r}
head(longData)
```

```{r}
library(gsheet)
dataUnido <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1uLaXke-KPN28-ESPPoihk8TiXVWp5xuNGHW7w7yqLCc/edit#gid=416085055")
dataUnido$isicCodeCombinaison <- NULL
dataUnido$tableDefinitionCode <- NULL
dataUnido$sourceCode <- NULL
dataUnido$updateYear <- NULL
dataUnido$unit <- NULL
```

# Dataset for dataCanada

```{r}
library(tidyverse)
dataCanada131 <- filter(dataUnido, countryCode == 76)
dataCanada131 <- filter(dataCanada131, isicCode == 131)
dataCanada131 <- filter(dataCanada131, tableCode == 4)
dataCanada131 <- filter(dataCanada131, year > 2008)
dataCanada131 <- dataCanada131 %>%
  pivot_wider(names_from = isicCode, values_from = value)
head(dataCanada131)
```

# Dataset for dataCanada181

```{r}
dataCanada181 <- filter(dataUnido, countryCode == 76)
dataCanada181 <- filter(dataCanada181, isicCode == 181)
dataCanada181 <- filter(dataCanada181, tableCode == 4)
dataCanada181 <- filter(dataCanada181, year > 2008)
dataCanada181 <- dataCanada181 %>%
  pivot_wider(names_from = isicCode, values_from = value)
head(dataCanada181)
```

# Merging 2 datasets

```{r}
dataCanadaFull <- left_join(dataCanada131, dataCanada181, c("year","tableCode","countryCode"))
head(dataCanadaFull)
```

# Transform dataCanadaFull in long data format

```{r}
dataCanadaFullLong <- dataCanadaFull %>% 
  pivot_longer(!c(year, tableCode, countryCode), names_to = "isicCode", values_to = "value")
```

